Machine Learning for Predictive Auto-Tuning

### 2011-12-12 paper/draft1.tex
Add some general HPC intro, moving to more motivation text

Current languages and software engineering practices have conspired to make it difficult
to provide high-performance libraries for general mathematical operations on GPU devices.
The dominant approach to implementing high-performance GPU codes has been auto-tuning,
which is fundamentally slow. Auto-tuning requires running a large number of
sub-compilations and representative computations to actually {\em measure} which
of several possible implementations is fastest. General-purpose libraries such
as CUBLAS and CUFFT XXX hide this overhead by auto-tuning offline and using
intimate knowledge of the mathematical problem being solved as well as the
kernel implementations to heuristically choose a good computation strategy quickly when a
user asks for e.g. a matrix product of particular arguments.
Designing and testing libraries of this form is time-consuming, difficult, and provides little
reusable infrastructure to assist in the implementation of different computations.
Libraries for stencil computations and sparse matrix operations must start from scratch.

>>>>

% HPC Landscape: from single proc to multi-cores

Due to power consumption and heat dissipation concerns, scientific applications
have shifted from computing platforms where performance had been primarily
driven by rises in the clock frequency of a single “heavy-weight” processor
(with complex out-of-order control and cache structures) to a platform with
ever increasing numbers of “light-weight” cores. Interestingly, this shift is
now not only relevant to computational sciences but to the development of all
computer systems: from ubiquitous consumer-facing devices (e.g. phones) to
high-end computer farms for web-scale applications (e.g. social networks).

% HPC Landscape: large diversity of architectures, no consensus => push to
% flexibility => push to sw

Even though the future lies in low-power multi-core hardware designs, the field
lacks meaningful consensus on exactly how the different subsystems (memory,
communication and computation) should be efficiently integrated, modelled and
programmed. As a consequence, these systems have been increasingly relying on
flexible but low-level software-controlled cache management and parallelism
\cite{BerkeleyViewOfHPC} in order to better control and understand the various
trade-offs among (e.g.) performance, reliability, energy efficiency, production
costs, etc. This evolution has profoundly altered the landscape of software
development: application programmers are now facing a large diversity of
low-level architectural features while being left with rapidly evolving tools
and techniques to write high-performance \emph{and} portable code.

% Better tools: motivation

% XXX

\subsection{Case Study}
Current languages and software engineering practices have conspired to make it difficult
to provide high-performance libraries for general mathematical operations on GPU devices.
The dominant approach to implementing high-performance GPU codes has been auto-tuning,
which is fundamentally slow. Auto-tuning requires running a large number of
sub-compilations and representative computations to actually {\em measure} which
of several possible implementations is fastest. General-purpose libraries such
as CUBLAS and CUFFT XXX hide this overhead by auto-tuning offline and using
intimate knowledge of the mathematical problem being solved as well as the
kernel implementations to heuristically choose a good computation strategy quickly when a
user asks for e.g. a matrix product of particular arguments.
Designing and testing libraries of this form is time-consuming, difficult, and provides little
reusable infrastructure to assist in the implementation of different computations.
Libraries for stencil computations and sparse matrix operations must start from scratch.
